The Journey from Wild to Textbook Data: A Case Study from the National Longitudinal Survey of Youth

### 2021-09-20 paper/paper.Rmd
Modify chunk label so that Figure 4 does not display same plot fig.2

Further, \ref{fig:comppict} A shows that after eliminating the extreme values, the highest value has decreased to be around $350. The spikes are still observed but not as extreme as the original data set. In Figure \ref{fig:comppict} B), we plot the three features of mean hourly wages, namely the minimum, median, and maximum value, transformed to log scale. The plot implies that the skewness is slightly negative. We also see that the three features are overlapped each other. It indicates that some ID's minimum wages are higher than some ID's maximum wages.

```{r comppict, fig.cap = "The distribution of the mean hourly wage after the extreme values are replaced. Plot A portrays the pattern of mean hourly wage of high school cohort from 1979 to 2018 of each ID in US Dollar; Plot B shows the distribution of their minimum, median, and maximum value transformed to log10 scale. We can see that some observations still had reasonbaly higher wages than the others. Also, some IDs' have a minimum wages that is higher than others' maximum wages.", fig.width=6, fig.height=4, eval=FALSE}
```

>>>>

Further, \ref{fig:fixed-feature-plot} A shows that after eliminating the extreme values, the highest value has decreased to be around $350. The spikes are still observed but not as extreme as the original data set. In Figure \ref{fixed-feature-plot} B, we plot the three features of mean hourly wages, namely the minimum, median, and maximum value, transformed to log scale. The plot implies that the skewness is slightly negative. We also see that the three features are overlapped each other. It indicates that some ID's minimum wages are higher than some ID's maximum wages.

### 2021-09-20 paper/paper.tex
Modify chunk label so that Figure 4 does not display same plot fig.2

% Pandoc citation processing
\newlength{\csllabelwidth}
\setlength{\csllabelwidth}{3em}
\newlength{\cslhangindent}
\setlength{\cslhangindent}{1.5em}
% for Pandoc 2.8 to 2.10.1
\newenvironment{cslreferences}%
  {}%
  {\par}
% For Pandoc 2.11+
\newenvironment{CSLReferences}[3] % #1 hanging-ident, #2 entry spacing
 {% don't indent paragraphs
  \setlength{\parindent}{0pt}
  % turn on hanging indent if param 1 is 1
  \ifodd #1 \everypar{\setlength{\hangindent}{\cslhangindent}}\ignorespaces\fi
  % set entry spacing
  \ifnum #2 > 0
  \setlength{\parskip}{#2\baselineskip}
  \fi
 }%
 {}
\usepackage{calc} % for calculating minipage widths
\newcommand{\CSLBlock}[1]{#1\hfill\break}
\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{#1}}
\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{#1}}
\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}

\usepackage{tcolorbox}
\newtcolorbox{tbox}{width=\textwidth,boxsep=5pt}
\usepackage{booktabs}
\usepackage{longtable}
\usepackage{array}
\usepackage{multirow}
\usepackage{wrapfig}
\usepackage{float}
\usepackage{colortbl}
\usepackage{pdflscape}
\usepackage{tabu}
\usepackage{threeparttable}
\usepackage{threeparttablex}
\usepackage[normalem]{ulem}
\usepackage{makecell}
\usepackage{xcolor}

@@

\textbf{Keywords}: Data cleaning; Data tidying; Longitudinal data; NLSY79; Open data; Initial data analysis; Outlier detection; Robust linear regression

@@

``Open data'' is data that is freely accessible, modifiable, and shareable by anyone for any purpose (Open Knowledge Foundation 2021). This type of data can be useful as example data in statistical textbooks and for research purposes. However, open data is often referred to as what we might call ``wild data'' because it requires substantial cleaning and tidying to tame it into textbook shape. Huebner Marianne, Vach, and Cessie (2016) emphasize that making the data cleaning process accountable and transparent is imperative. Documenting the data cleaning is essential for the integrity of downstream statistical analyses and model building (M. Huebner et al. 2020).
Data cleaning can be considered to be a part of what is called ``initial data analysis (IDA)'' (Chatfield 1985). In IDA one would also explore the data, especially to check if the dta is consistent with assumptions required for modeling. This is also related to exploratory data analysis, which Tukey (1977) defines as detective work to learn from data, either numerically or graphically, before confirmatory data analysis is performed, encompasses IDA. Dasu and Johnson (2003) say that data cleaning and exploration, without naming it as IDA, is a difficult task and consumes 80\% of the data mining task.

Despite its importance, this IDA stage is often undervalued and neglected (Chatfield 1985). There are few research papers that document the data cleaning (Wickham 2014). Furthermore, the decisions made in this stage often go unreported in the sense that IDA is often performed in an unplanned and unstructured way and is only shared among restricted parties (M. Huebner et al. 2020).

This paper aims to demonstrate the steps of IDA, tidying, cleaning, and documenting the process, for a prominent open data source, National Longitudinal Survey of Youth (NLSY) 1979, henceforth referred to NLSY79. This data has been playing an important role in research in various disciplines, including but are not limited to economics, sociology, education, public policy, and public health for more than a quarter of the century (Pergamit et al. 2001). In addition, The National Longitudinal Survey is considered a survey with high retention rates and carefully designed, making it suitable for a life course research (Pergamit et al. 2001) and (Cooksey 2017). According to Cooksey (2017), thousand of articles, and hundreds of book chapters and monographs has incorporated the NLSY data. Moreover, the NLSY79 is considered the most widely used and most important cohort in the NLSY79 data sets (Pergamit et al. 2001).

@@

@@

Education, Training and Achievement Scores

\begin{itemize}
\tightlist
\item
  Education -\textgreater{} Summary measures -\textgreater{} All schools -\textgreater{} By year -\textgreater{} Highest grade completed

  \begin{itemize}
  \tightlist
  \item
    Downloaded all of variables in Highest grade completed.
  \end{itemize}
\end{itemize}

Employment

\begin{itemize}
\tightlist
\item
  Summary measures -\textgreater{} By job -\textgreater{} Hours worked and Hourly wages

  \begin{itemize}
  \tightlist
  \item
    Downloaded all of the variables in Hours worked.
  \item
    Downloaded all of the variables in Hourly wages
    Both hours worked and hourly wages are recorded by the job, up to five jobs for each id/subject.
  \end{itemize}
\end{itemize}

Household, Geography and Contextual Variables

\begin{itemize}
\tightlist
\item
  Context -\textgreater{} Summary measures -\textgreater{} Basic demographics

  \begin{itemize}
  \tightlist
  \item
    Downloaded year and month of birth, race, and sex variables.
    There are two versions of the year and month of birth, i.e., 1979 and 1981 data. We downloaded these two versions.
  \end{itemize}
\end{itemize}

@@

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{get\_rate }\OtherTok{\textless{}{-}} \ControlFlowTok{function}\NormalTok{(year) \{}
\NormalTok{  new\_data\_qnames }\SpecialCharTok{\%\textgreater{}\%}
\NormalTok{    dplyr}\SpecialCharTok{::}\FunctionTok{select}\NormalTok{(CASEID\_1979,}
                  \FunctionTok{starts\_with}\NormalTok{(}\StringTok{"HRP"}\NormalTok{) }\SpecialCharTok{\&}
                    \FunctionTok{ends\_with}\NormalTok{(}\FunctionTok{as.character}\NormalTok{(year))) }\SpecialCharTok{\%\textgreater{}\%}
    \FunctionTok{pivot\_longer}\NormalTok{(}\SpecialCharTok{!}\NormalTok{CASEID\_1979, }\AttributeTok{names\_to =} \StringTok{"job"}\NormalTok{, }\AttributeTok{values\_to =} \StringTok{"rate\_per\_hour"}\NormalTok{) }\SpecialCharTok{\%\textgreater{}\%}
    \FunctionTok{separate}\NormalTok{(}\StringTok{"job"}\NormalTok{, }\FunctionTok{c}\NormalTok{(}\StringTok{"job"}\NormalTok{, }\StringTok{"year"}\NormalTok{), }\AttributeTok{sep =} \SpecialCharTok{{-}}\DecValTok{4}\NormalTok{) }\SpecialCharTok{\%\textgreater{}\%}
    \FunctionTok{mutate}\NormalTok{(}\AttributeTok{job =} \FunctionTok{paste0}\NormalTok{(}\StringTok{"job\_0"}\NormalTok{, }\FunctionTok{substr}\NormalTok{(job, }\DecValTok{4}\NormalTok{, }\DecValTok{4}\NormalTok{))) }\SpecialCharTok{\%\textgreater{}\%}
    \FunctionTok{rename}\NormalTok{(}\AttributeTok{id =}\NormalTok{ CASEID\_1979)}
\NormalTok{\}}
\NormalTok{rates }\OtherTok{\textless{}{-}} \FunctionTok{list}\NormalTok{()}
\ControlFlowTok{for}\NormalTok{(ayear }\ControlFlowTok{in} \FunctionTok{c}\NormalTok{(}\DecValTok{1979}\SpecialCharTok{:}\DecValTok{1994}\NormalTok{, }\DecValTok{1996}\NormalTok{, }\DecValTok{1998}\NormalTok{, }\DecValTok{2000}\NormalTok{, }\DecValTok{2002}\NormalTok{, }\DecValTok{2004}\NormalTok{, }\DecValTok{2006}\NormalTok{, }\DecValTok{2008}\NormalTok{, }\DecValTok{2010}\NormalTok{,}
               \DecValTok{2012}\NormalTok{, }\DecValTok{2014}\NormalTok{, }\DecValTok{2016}\NormalTok{, }\DecValTok{2018}\NormalTok{)) \{}
\NormalTok{  rates[[ayear]] }\OtherTok{\textless{}{-}} \FunctionTok{get\_rate}\NormalTok{(ayear)}
\NormalTok{\}}
\NormalTok{rates\_all }\OtherTok{\textless{}{-}} \FunctionTok{bind\_rows}\NormalTok{(}\SpecialCharTok{!!!}\NormalTok{rates)}
\CommentTok{\# join hours and rates variable}
\NormalTok{hours\_wages }\OtherTok{\textless{}{-}} \FunctionTok{left\_join}\NormalTok{(rates\_all,}
\NormalTok{                         hours\_all,}
                         \AttributeTok{by =} \FunctionTok{c}\NormalTok{(}\StringTok{"id"}\NormalTok{, }\StringTok{"year"}\NormalTok{, }\StringTok{"job"}\NormalTok{)) }\SpecialCharTok{\%\textgreater{}\%}
  \CommentTok{\# set the 0 value in rate\_per\_hour as NA}
  \FunctionTok{mutate}\NormalTok{(}\AttributeTok{rate\_per\_hour =} \FunctionTok{ifelse}\NormalTok{(rate\_per\_hour }\SpecialCharTok{==} \DecValTok{0}\NormalTok{, }\ConstantTok{NA}\NormalTok{,}
\NormalTok{                                rate\_per\_hour),}
         \AttributeTok{hours\_work =} \FunctionTok{ifelse}\NormalTok{(hours\_work }\SpecialCharTok{==} \DecValTok{0} \SpecialCharTok{|}\NormalTok{ hours\_work }\SpecialCharTok{\textgreater{}} \DecValTok{84}\NormalTok{, }\ConstantTok{NA}\NormalTok{, hours\_work))}
\end{Highlighting}
\end{Shaded}

@@

Further, \ref{fig:comppict} A shows that after eliminating the extreme values, the highest value has decreased to be around \$350. The spikes are still observed but not as extreme as the original data set. In Figure \ref{fig:comppict} B), we plot the three features of mean hourly wages, namely the minimum, median, and maximum value, transformed to log scale. The plot implies that the skewness is slightly negative. We also see that the three features are overlapped each other. It indicates that some ID's minimum wages are higher than some ID's maximum wages.

@@

The entire analysis is conducted using \texttt{R} (R Core Team 2020) in \texttt{rstudio} using these packages: \texttt{tidyverse} (Wickham et al. 2019), \texttt{ggplot2} (Wickham 2016), \texttt{dplyr} (Wickham et al. 2020), \texttt{readr} (Wickham and Hester 2020), \texttt{tidyr} (Wickham 2020b), \texttt{stringr} (Wickham 2019), \texttt{purrr} (Henry and Wickham 2020), \texttt{broom} (Robinson, Hayes, and Couch 2021), \texttt{brolgar} (Tierney, Cook, and Prvan 2020), \texttt{patchwork} (Pedersen 2020), \texttt{kableExtra} (Zhu 2019), \texttt{MASS} (Venables and Ripley 2002), \texttt{janitor} (Firke 2020), \texttt{DiagrammeR} (Iannone 2020), \texttt{rsvg} (Ooms 2020), \texttt{webshot} (Chang 2019), \texttt{mgcv} (Wood 2003), \texttt{tsibble} (Wang, Cook, and Hyndman 2020), and \texttt{modelr} (Wickham 2020a). The paper are generated using \texttt{knitr} (Xie 2014), \texttt{rmarkdown} (Xie, Dervieux, and Riederer 2020), and \texttt{rticles} (Allaire et al. 2021).

@@

\hypertarget{refs}{}
\begin{CSLReferences}{1}{0}
\leavevmode\hypertarget{ref-rticles}{}%
Allaire, JJ, Yihui Xie, R Foundation, Hadley Wickham, Journal of Statistical Software, Ramnath Vaidyanathan, Association for Computing Machinery, et al. 2021. \emph{Rticles: Article Formats for r Markdown}. \url{https://CRAN.R-project.org/package=rticles}.

@@

\leavevmode\hypertarget{ref-shiny}{}%
Chang, Winston, Joe Cheng, JJ Allaire, Yihui Xie, and Jonathan McPherson. 2020. \emph{Shiny: Web Application Framework for r}. \url{https://CRAN.R-project.org/package=shiny}.

\leavevmode\hypertarget{ref-Chatfield1985TIEo}{}%
Chatfield, C. 1985. {``The Initial Examination of Data.''} \emph{Journal of the Royal Statistical Society. Series A. General} 148 (3): 214--53.

\leavevmode\hypertarget{ref-eliznlsy}{}%
Cooksey, Elizabeth C. 2017. {``Using the National Longitudinal Surveys of Youth (NLSY) to Conduct Life Course Analyses.''} In \emph{Handbook of Life Course Health Development}, edited by Richard M. Lerner Neal Halfon Christoper B. Forrest, 561--77. Cham: Springer. https://doi.org/\url{https://doi.org/10.1007/978-3-319-47143-3_23}.

@@

\leavevmode\hypertarget{ref-HuebnerMariannePhD2016Asat}{}%
Huebner, Marianne, Werner Vach, and Saskia le Cessie. 2016. {``A Systematic Approach to Initial Data Analysis Is Good Research Practice.''} \emph{The Journal of Thoracic and Cardiovascular Surgery} 151 (1): 25--27.

\leavevmode\hypertarget{ref-HuebnerMarianne2020Haar}{}%
Huebner, Marianne, Werner Vach, Saskia le Cessie, Carsten Oliver Schmidt, and Lara Lusa. 2020. {``Hidden Analyses: A Review of Reporting Practice and Recommendations for More Transparent Reporting of Initial Data Analyses.''} \emph{BMC Medical Research Methodology} 20 (1): 61--61.

@@

\leavevmode\hypertarget{ref-KollerManuel2016rARP}{}%
Koller, Manuel. 2016. {``Robustlmm: An r Package for Robust Estimation of Linear Mixed-Effects Models.''} \emph{Journal of Statistical Software} 75 (6): 1--24.

\leavevmode\hypertarget{ref-OECD}{}%
OECD. n.d. {``Gender Wage Gap.''} https://data.oecd.org/earnwage/gender-wage-gap.htm.

\leavevmode\hypertarget{ref-rsvg}{}%
Ooms, Jeroen. 2020. \emph{Rsvg: Render SVG Images into PDF, PNG, PostScript, or Bitmap Arrays}. \url{https://CRAN.R-project.org/package=rsvg}.

\leavevmode\hypertarget{ref-opendata}{}%
Open Knowledge Foundation. 2021. {``Open Definition. Defining Open in Open Data, Open Content, and Open Knowledge.''} 2021. \url{http://opendefinition.org/od/2.1/en/}.

@@

\leavevmode\hypertarget{ref-MichaelRPergamit2001DWTN}{}%
Pergamit, Michael R., Charles R. Pierret, Donna S. Rothstein, and Jonathan R. Veum. 2001. {``Data Watch: The National Longitudinal Surveys.''} \emph{The Journal of Economic Perspectives} 15 (2): 239--53.

@@

\leavevmode\hypertarget{ref-nlsy79}{}%
The U.S. Bureau of Labor Statistics. n.d. {``National Longitudinal Survey of Youth 1979.''} Available at \url{https://www.nlsinfo.org/content/cohorts/nlsy79} (2021/25/02).

\leavevmode\hypertarget{ref-brolgar}{}%
Tierney, Nicholas, Di Cook, and Tania Prvan. 2020. \emph{Brolgar: BRowse over Longitudinal Data Graphically and Analytically in r}. \url{https://github.com/njtierney/brolgar}.

@@

\leavevmode\hypertarget{ref-rlm}{}%
UCLA: Statistical Consulting Group. 2021. {``Robust Regression \textbar{} r Data Analysis Examples.''} February 2021. \url{https://stats.idre.ucla.edu/r/dae/robust-regression/}.

\leavevmode\hypertarget{ref-validate}{}%
van der Loo, Mark P. J., and Edwin de Jonge. 2021. {``Data Validation Infrastructure for {R}.''} \emph{Journal of Statistical Software} 97 (10): 1--31. \url{https://doi.org/10.18637/jss.v097.i10}.

\leavevmode\hypertarget{ref-mass}{}%
Venables, W. N., and B. D. Ripley. 2002. \emph{Modern Applied Statistics with s}. Fourth. New York: Springer. \url{http://www.stats.ox.ac.uk/pub/MASS4}.

\leavevmode\hypertarget{ref-tsibble}{}%
Wang, Earo, Dianne Cook, and Rob J Hyndman. 2020. {``A New Tidy Data Structure to Support Exploration and Modeling of Temporal Data.''} \emph{Journal of Computational and Graphical Statistics} 29 (3): 466--78. \url{https://doi.org/10.1080/10618600.2019.1695624}.

\leavevmode\hypertarget{ref-WickhamHadley2014TD}{}%
Wickham, Hadley. 2014. {``Tidy Data.''} \emph{Journal of Statistical Software} 59 (10): 1--23.

@@

\leavevmode\hypertarget{ref-tidyverse}{}%
Wickham, Hadley, Mara Averick, Jennifer Bryan, Winston Chang, Lucy D'Agostino McGowan, Romain François, Garrett Grolemund, et al. 2019. {``Welcome to the {tidyverse}.''} \emph{Journal of Open Source Software} 4 (43): 1686. \url{https://doi.org/10.21105/joss.01686}.

@@

\leavevmode\hypertarget{ref-kableExtra}{}%
Zhu, Hao. 2019. \emph{kableExtra: Construct Complex Table with 'Kable' and Pipe Syntax}. \url{https://CRAN.R-project.org/package=kableExtra}.

\end{CSLReferences}

>>>>

% Pandoc citation processing

\usepackage{tcolorbox}
\usepackage{fontawesome}
\usepackage{booktabs}
\usepackage{longtable}
\usepackage{array}
\usepackage{multirow}
\usepackage{wrapfig}
\usepackage{float}
\usepackage{colortbl}
\usepackage{pdflscape}
\usepackage{tabu}
\usepackage{threeparttable}
\usepackage{threeparttablex}
\usepackage[normalem]{ulem}
\usepackage{makecell}
\usepackage{xcolor}

@@

\keywords{
    Data cleaning; Data tidying; Longitudinal data; NLSY79; Open data; Initial data analysis; Outlier detection; Robust linear regression
  }

@@

``Open data'' is data that is freely accessible, modifiable, and shareable by anyone for any purpose (Open Knowledge Foundation 2021). This type of data can be useful as example data in statistical textbooks and for research purposes. However, open data is often referred to as what we might call ``wild data'' because it requires substantial cleaning and tidying to tame it into textbook shape. Huebner, Vach, and Cessie (2016) emphasize that making the data cleaning process accountable and transparent is imperative. Documenting the data cleaning is essential for the integrity of downstream statistical analyses and model building (Huebner et al. 2020).
Data cleaning can be considered to be a part of what is called ``initial data analysis (IDA)'' (Chatfield 1985). In IDA one would also explore the data, especially to check if the data is consistent with assumptions required for modeling. This is also related to exploratory data analysis, which Tukey (1977) defines as detective work to learn from data, either numerically or graphically, before confirmatory data analysis is performed, which encompasses IDA. Dasu and Johnson (2003) say that data cleaning and exploration, without naming it as IDA, is a difficult task and consumes 80\% of the data mining task.

Despite its importance, this IDA stage is often undervalued and neglected (Chatfield 1985). There are few research papers that document the data cleaning (Wickham 2014). Furthermore, the decisions made in this stage often go unreported in the sense that IDA is often performed in an unplanned and unstructured way and is only shared among restricted parties (Huebner et al. 2020).

This paper aims to demonstrate the steps of IDA, tidying, cleaning, and documenting the process, for a prominent open data source, National Longitudinal Survey of Youth (NLSY) 1979, henceforth referred to NLSY79. This data has been playing an important role in research in various disciplines, including but are not limited to economics, sociology, education, public policy, and public health for more than a quarter of the century (Pergamit et al. 2001). In addition, The NLSY is considered a survey with high retention rates and carefully designed, making it suitable for a life course research (Pergamit et al. 2001; Cooksey 2017). According to Cooksey (2017), thousand of articles, and hundreds of book chapters and monographs has incorporated the NLSY data. Moreover, the NLSY79 is considered the most widely used and most important cohort in the NLSY79 data sets (Pergamit et al. 2001).

@@

In this section, we outline the steps taken to the download the raw data from the database provided at \url{https://www.nlsinfo.org/content/cohorts/nlsy79/get-data} (The U.S. Bureau of Labor Statistics, n.d.). Specifically, we only extract a subset of the variables in the database as shown in \ref{box:data}.

@@

\begin{tcolorbox}[label="box:data", title=Getting the data]

\faDatabase National Longitudinal Survey of Youth 1979

\begin{itemize}
\item[$\triangleright$] Education, Training and Achievement Scores
\begin{itemize}
\item[$\triangleright$] Education $\triangleright$ Summary measures $\triangleright$ All schools $\triangleright$ By year $\triangleright$ Highest grade completed
\begin{itemize}
\item[\faCloudDownload] Downloaded all of variables in Highest grade completed.
\end{itemize}
\end{itemize}
\item[$\triangleright$] Employment
\begin{itemize}
\item[$\triangleright$] Summary measures $\triangleright$ By job $\triangleright$ Hours worked and Hourly wages
\begin{itemize}
\item[\faCloudDownload] Downloaded all of the variables in Hours worked.
\item[\faCloudDownload] Downloaded all of the variables in Hourly wages
  Both hours worked and hourly wages are recorded by the job, up to five jobs for each id/subject. 
\end{itemize}
\end{itemize}
\item[$\triangleright$] Household, Geography and Contextual Variables
\begin{itemize}
\item[$\triangleright$] Context $\triangleright$ Summary measures $\triangleright$ Basic demographics 
\begin{itemize}
\item[\faCloudDownload] Downloaded year and month of birth, race, and sex variables. There are two versions of the year and month of birth, i.e., 1979 and 1981 data. We downloaded these two versions.
\end{itemize}
\end{itemize}
\end{itemize}

\end{tcolorbox}

@@

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{get_rate <-}\StringTok{ }\ControlFlowTok{function}\NormalTok{(year) \{}
\NormalTok{  new_data_qnames }\OperatorTok{%>%}
\StringTok{    }\NormalTok{dplyr}\OperatorTok{::}\KeywordTok{select}\NormalTok{(CASEID_}\DecValTok{1979}\NormalTok{,}
                  \KeywordTok{starts_with}\NormalTok{(}\StringTok{"HRP"}\NormalTok{) }\OperatorTok{&}
\StringTok{                    }\KeywordTok{ends_with}\NormalTok{(}\KeywordTok{as.character}\NormalTok{(year))) }\OperatorTok{%>%}
\StringTok{    }\KeywordTok{pivot_longer}\NormalTok{(}\OperatorTok{!}\NormalTok{CASEID_}\DecValTok{1979}\NormalTok{, }\DataTypeTok{names_to =} \StringTok{"job"}\NormalTok{, }\DataTypeTok{values_to =} \StringTok{"rate_per_hour"}\NormalTok{) }\OperatorTok{%>%}
\StringTok{    }\KeywordTok{separate}\NormalTok{(}\StringTok{"job"}\NormalTok{, }\KeywordTok{c}\NormalTok{(}\StringTok{"job"}\NormalTok{, }\StringTok{"year"}\NormalTok{), }\DataTypeTok{sep =} \DecValTok{-4}\NormalTok{) }\OperatorTok{%>%}
\StringTok{    }\KeywordTok{mutate}\NormalTok{(}\DataTypeTok{job =} \KeywordTok{paste0}\NormalTok{(}\StringTok{"job_0"}\NormalTok{, }\KeywordTok{substr}\NormalTok{(job, }\DecValTok{4}\NormalTok{, }\DecValTok{4}\NormalTok{))) }\OperatorTok{%>%}
\StringTok{    }\KeywordTok{rename}\NormalTok{(}\DataTypeTok{id =}\NormalTok{ CASEID_}\DecValTok{1979}\NormalTok{)}
\NormalTok{\}}
\NormalTok{rates <-}\StringTok{ }\KeywordTok{list}\NormalTok{()}
\ControlFlowTok{for}\NormalTok{(ayear }\ControlFlowTok{in} \KeywordTok{c}\NormalTok{(}\DecValTok{1979}\OperatorTok{:}\DecValTok{1994}\NormalTok{, }\DecValTok{1996}\NormalTok{, }\DecValTok{1998}\NormalTok{, }\DecValTok{2000}\NormalTok{, }\DecValTok{2002}\NormalTok{, }\DecValTok{2004}\NormalTok{, }\DecValTok{2006}\NormalTok{, }\DecValTok{2008}\NormalTok{, }\DecValTok{2010}\NormalTok{,}
               \DecValTok{2012}\NormalTok{, }\DecValTok{2014}\NormalTok{, }\DecValTok{2016}\NormalTok{, }\DecValTok{2018}\NormalTok{)) \{}
\NormalTok{  rates[[ayear]] <-}\StringTok{ }\KeywordTok{get_rate}\NormalTok{(ayear)}
\NormalTok{\}}
\NormalTok{rates_all <-}\StringTok{ }\KeywordTok{bind_rows}\NormalTok{(}\OperatorTok{!!!}\NormalTok{rates)}
\CommentTok{# join hours and rates variable}
\NormalTok{hours_wages <-}\StringTok{ }\KeywordTok{left_join}\NormalTok{(rates_all,}
\NormalTok{                         hours_all,}
                         \DataTypeTok{by =} \KeywordTok{c}\NormalTok{(}\StringTok{"id"}\NormalTok{, }\StringTok{"year"}\NormalTok{, }\StringTok{"job"}\NormalTok{)) }\OperatorTok{%>%}
\StringTok{  }\CommentTok{# set the 0 value in rate_per_hour as NA}
\StringTok{  }\KeywordTok{mutate}\NormalTok{(}\DataTypeTok{rate_per_hour =} \KeywordTok{ifelse}\NormalTok{(rate_per_hour }\OperatorTok{==}\StringTok{ }\DecValTok{0}\NormalTok{, }\OtherTok{NA}\NormalTok{,}
\NormalTok{                                rate_per_hour),}
         \DataTypeTok{hours_work =} \KeywordTok{ifelse}\NormalTok{(hours_work }\OperatorTok{==}\StringTok{ }\DecValTok{0} \OperatorTok{|}\StringTok{ }\NormalTok{hours_work }\OperatorTok{>}\StringTok{ }\DecValTok{84}\NormalTok{, }\OtherTok{NA}\NormalTok{, hours_work))}
\end{Highlighting}
\end{Shaded}

@@

Further, \ref{fig:fixed-feature-plot} A shows that after eliminating the extreme values, the highest value has decreased to be around \$350. The spikes are still observed but not as extreme as the original data set. In Figure \ref{fixed-feature-plot} B, we plot the three features of mean hourly wages, namely the minimum, median, and maximum value, transformed to log scale. The plot implies that the skewness is slightly negative. We also see that the three features are overlapped each other. It indicates that some ID's minimum wages are higher than some ID's maximum wages.

@@

The entire analysis is conducted using \texttt{R} (R Core Team 2020) in \texttt{rstudio} using these packages: \texttt{tidyverse} (H. Wickham, Averick, et al. 2019), \texttt{ggplot2} (Wickham 2016), \texttt{dplyr} (H. Wickham, François, et al. 2020), \texttt{readr} (H. Wickham and Hester 2020), \texttt{tidyr} (H. Wickham 2020b), \texttt{stringr} (H. Wickham 2019), \texttt{purrr} (Henry and Wickham 2020), \texttt{broom} (Robinson, Hayes, and Couch 2021), \texttt{brolgar} (Tierney, Cook, and Prvan 2020), \texttt{patchwork} (Pedersen 2020), \texttt{kableExtra} (Zhu 2019), \texttt{MASS} (Venables and Ripley 2002), \texttt{janitor} (Firke 2020), \texttt{DiagrammeR} (Iannone 2020), \texttt{rsvg} (Ooms 2020), \texttt{webshot} (Chang 2019), \texttt{mgcv} (Wood 2003), \texttt{tsibble} (Wang, Cook, and Hyndman 2020), and \texttt{modelr} (H. Wickham 2020a). The paper are generated using \texttt{knitr} (Xie 2014), \texttt{rmarkdown} (Xie, Dervieux, and Riederer 2020), and \texttt{rticles} (Allaire et al. 2021).

@@

\hypertarget{refs}{}
\leavevmode\hypertarget{ref-rticles}{}%
Allaire, JJ, Yihui Xie, R Foundation, Hadley Wickham, Journal of Statistical Software, Ramnath Vaidyanathan, Association for Computing Machinery, et al. 2021. \emph{Rticles: Article Formats for R Markdown}. \url{https://CRAN.R-project.org/package=rticles}.

@@

\leavevmode\hypertarget{ref-shiny}{}%
Chang, Winston, Joe Cheng, JJ Allaire, Yihui Xie, and Jonathan McPherson. 2020. \emph{Shiny: Web Application Framework for R}. \url{https://CRAN.R-project.org/package=shiny}.

\leavevmode\hypertarget{ref-Chatfield1985TIEo}{}%
Chatfield, C. 1985. ``The Initial Examination of Data.'' \emph{Journal of the Royal Statistical Society. Series A. General} 148 (3): 214--53.

\leavevmode\hypertarget{ref-eliznlsy}{}%
Cooksey, Elizabeth C. 2017. ``Using the National Longitudinal Surveys of Youth (Nlsy) to Conduct Life Course Analyses.'' In \emph{Handbook of Life Course Health Development}, edited by Richard M. Lerner Neal Halfon Christoper B. Forrest, 561--77. Cham: Springer. \url{https://doi.org/https://doi.org/10.1007/978-3-319-47143-3_23}.

@@

\leavevmode\hypertarget{ref-HuebnerMariannePhD2016Asat}{}%
Huebner, Marianne, Werner Vach, and Saskia le Cessie. 2016. ``A Systematic Approach to Initial Data Analysis Is Good Research Practice.'' \emph{The Journal of Thoracic and Cardiovascular Surgery} 151 (1): 25--27.

\leavevmode\hypertarget{ref-HuebnerMarianne2020Haar}{}%
Huebner, Marianne, Werner Vach, Saskia le Cessie, Carsten Oliver Schmidt, and Lara Lusa. 2020. ``Hidden Analyses: A Review of Reporting Practice and Recommendations for More Transparent Reporting of Initial Data Analyses.'' \emph{BMC Medical Research Methodology} 20 (1): 61--61.

@@

\leavevmode\hypertarget{ref-KollerManuel2016rARP}{}%
Koller, Manuel. 2016. ``Robustlmm: An R Package for Robust Estimation of Linear Mixed-Effects Models.'' \emph{Journal of Statistical Software} 75 (6): 1--24.

\leavevmode\hypertarget{ref-OECD}{}%
OECD. n.d. ``Gender Wage Gap.'' https://data.oecd.org/earnwage/gender-wage-gap.htm.

\leavevmode\hypertarget{ref-rsvg}{}%
Ooms, Jeroen. 2020. \emph{Rsvg: Render Svg Images into Pdf, Png, Postscript, or Bitmap Arrays}. \url{https://CRAN.R-project.org/package=rsvg}.

\leavevmode\hypertarget{ref-opendata}{}%
Open Knowledge Foundation. 2021. ``Open Definition. Defining Open in Open Data, Open Content, and Open Knowledge.'' 2021. \url{http://opendefinition.org/od/2.1/en/}.

@@

\leavevmode\hypertarget{ref-MichaelRPergamit2001DWTN}{}%
Pergamit, Michael R., Charles R. Pierret, Donna S. Rothstein, and Jonathan R. Veum. 2001. ``Data Watch: The National Longitudinal Surveys.'' \emph{The Journal of Economic Perspectives} 15 (2): 239--53.

@@

\leavevmode\hypertarget{ref-nlsy79}{}%
The U.S. Bureau of Labor Statistics. n.d. ``National Longitudinal Survey of Youth 1979.'' Available at \url{https://www.nlsinfo.org/content/cohorts/nlsy79} (2021/25/02).

\leavevmode\hypertarget{ref-brolgar}{}%
Tierney, Nicholas, Di Cook, and Tania Prvan. 2020. \emph{Brolgar: BRowse over Longitudinal Data Graphically and Analytically in R}. \url{https://github.com/njtierney/brolgar}.

@@

\leavevmode\hypertarget{ref-rlm}{}%
UCLA: Statistical Consulting Group. 2021. ``Robust Regression \textbar{} R Data Analysis Examples.'' February 2021. \url{https://stats.idre.ucla.edu/r/dae/robust-regression/}.

\leavevmode\hypertarget{ref-validate}{}%
van der Loo, Mark P. J., and Edwin de Jonge. 2021. ``Data Validation Infrastructure for R.'' \emph{Journal of Statistical Software} 97 (10): 1--31. \url{https://doi.org/10.18637/jss.v097.i10}.

\leavevmode\hypertarget{ref-mass}{}%
Venables, W. N., and B. D. Ripley. 2002. \emph{Modern Applied Statistics with S}. Fourth. New York: Springer. \url{http://www.stats.ox.ac.uk/pub/MASS4}.

\leavevmode\hypertarget{ref-tsibble}{}%
Wang, Earo, Dianne Cook, and Rob J Hyndman. 2020. ``A New Tidy Data Structure to Support Exploration and Modeling of Temporal Data.'' \emph{Journal of Computational and Graphical Statistics} 29 (3): 466--78. \url{https://doi.org/10.1080/10618600.2019.1695624}.

\leavevmode\hypertarget{ref-WickhamHadley2014TD}{}%
Wickham, Hadley. 2014. ``Tidy Data.'' \emph{Journal of Statistical Software} 59 (10): 1--23.

@@

\leavevmode\hypertarget{ref-tidyverse}{}%
Wickham, Hadley, Mara Averick, Jennifer Bryan, Winston Chang, Lucy D'Agostino McGowan, Romain François, Garrett Grolemund, et al. 2019. ``Welcome to the tidyverse.'' \emph{Journal of Open Source Software} 4 (43): 1686. \url{https://doi.org/10.21105/joss.01686}.

@@

\leavevmode\hypertarget{ref-kableExtra}{}%
Zhu, Hao. 2019. \emph{KableExtra: Construct Complex Table with 'Kable' and Pipe Syntax}. \url{https://CRAN.R-project.org/package=kableExtra}.
